### 2025-06-01 conteudo/results/students.tex
fix grammar with grammarly

\section{Students}

As presented in chapter \ref{cha:method}, we propose to students to mitigate duplications
found by the ArKanjo tool and send patches to the linux subsystems maintainers, while
documenting their experiences and learning through the process.

For the students that worked
in the IIO subsytems, we gave them duplicated function pairs found by the tool and their 
responsability was to think of how to mitigate the duplication and sent the patch to the 
maintainers. The student that worked in the AMD Display driver have the additional work
to use the tool to find the duplications.

Appendix \ref{app:stuart} has references
to the artifacts discussed in this section, including the duplication function pairs that students 
mitigate, the students's reports, the patch students sent to the linux kernel, and the 
answers of the questionary asked to the student that worked in the AMD Display driver and the 
teaching assistant that used the tool to find duplications in the IIO.

@@

Table \ref{tab:stu} summarizes the mitigation approachs of students's groups in terms 
of lines changed and refactor methods used. A total of seven groups applied the parameterize refactor method,
and two groups applied the extractor method on their approaches to mitigate the duplications. 
The other three groups does not applied any refactor methods presented in the literature. 
We discuss the groups's approaches and the subsystem maintainers opinions on the patches that students
sent.

Groups one and four have worked in similar duplications, where there are two functions returning a boolean mirroring
the results. That means, when the first function returns true, the second function returns false, and when the first
functions returns false, the second function return false. The group one have a simple approach and the maintainers 
approved the changes quickly. The group four attempted to mitigate the duplication with macros, which is a method 
found in the code file they were working. The maintainers disagree with the proposed approach, arguing 
that the proposed changes decreases the code readability and asked for a approach similar to the group one. 
The group four submitted the patch after coding the approach proposed by the maintainers.

Group three, six and ten created a struct to encapsulate the parameters to mitigate the duplication when 
applying the parameterize method. The group three mitigation was send to the maintainers and the approved 
without significant comments, while the patch from group ten was reject by the same maintainer that approved
the patch from group three. The maintainer justification to reject the patch is that group ten refactored 
code files that has simple context, threfore adding a layer of abstraction to remove a duplication 
deacrease the code readability without clear gains on the codebase quality.

Group seven and eight used the extract method to approach the mitigations. The group eight send the 
mitigations to the maintainers and the code changes was approved with minor requested changes
about code style and smaller errors fix. For the group seven mitigation, the maintainers reject 
their patch because the code changes adds layers of abstractions on the code, turning it more complex
and harder to read. The maintainers proposed to the group seven to remove the duplicated functions 
and insert the duplicated code multiple times were the functions were being called. 
This suggestion we could think that it is a bad practice for code quality, as adds more complexity and 
responsibality to a bigger function, it is view as the best choice for removing layers of abstraction 
and making the code easier to read.

Group nine and eleven faced similar issue as us, where configurations files
designed as macros complicates to approach the mitigations. Group nine applied the parameterize method
to not need to know the configurations in the generic function, which is reasonable as the duplication
approach only depends on one configuration value. Group eleven worked with multiple configuration values,
and to refactor the duplication they proposed to create a bigger macro that encapsulates the other macros,
which is a clever solution but adds abstraction and complexity on macros, which makes the code harder
to debug and read. We do not get the feedback of the maintainers on the group eleven approach to know their 
opinions.

Group twelve is the student that executed the tool and pass throught the process of findings duplications.
The student found a duplication in the driver that happens accross eighteen code files, creating a patch
of removing four hundreds lines of duplicated code. 
The student send the mitigations to the maintainers and their feedback pointed to fix warning in the code, 
without arguing about removing the duplications.

Analysing the students approaches for the duplications and the maintainers feedback, 
we saw that people without previous experience in the linux kernel development could approach 
duplications found by the ArKanjo tool and become contributors to the kernel. We observed that 
not all duplications are view as code with bad quality, with the maintainers analysing the trade off
between what is the purpose of the duplicated code, the code readability, and the levels of abstraction 
added to address the duplications.

\subsection{Questionary Answers}

BLABLABLA

>>>>

\section{Study with University Students}

As presented in Chapter \ref{cha:method}, we propose that students mitigate duplications 
found by the ArKanjo tool, send patches to the Linux subsystem maintainers, and document 
their experiences and learnings.

For the students who worked in the IIO subsystems, we gave them duplicated function pairs 
found by the tool, and their responsibility was to think of how to mitigate the duplication 
and send the patch to the maintainers. The student that worked in the AMD Display driver 
have the additional work to use the tool to find the duplications.

Appendix \ref{app:stuart} has references to the artifacts discussed in this section, 
including the duplication function pairs that students mitigate, the students’ reports, 
the patch students sent to the Linux kernel, and the answers to the questionnaire asked 
of the students who worked in the AMD Display driver and the teaching assistant that used 
the tool to find duplications in the IIO.

@@

Table \ref{tab:stu} summarizes students' groups' mitigation approaches regarding lines 
changed and refactor methods used. Seven groups applied the parameterized refactor method, 
and two applied the extractor method to their approaches to mitigate the duplications. 
The other three groups do not use any refactor methods presented in the literature. 
We discuss the groups' approaches and the subsystem maintainers' opinions on the patches 
that students sent.

Groups one and four have worked in similar duplications, where two functions return a 
boolean mirroring the results. That means when the first function returns true, the second 
function returns false, and when the first function returns false, the second function 
returns false. Group one had a simple approach, and the maintainers approved the changes quickly. 
Group four attempted to mitigate the duplication with macros, a method found in the code file 
they were working on. The maintainers disagreed with the proposed approach, arguing that the 
proposed changes decrease the code readability, and asked for an approach similar to group one. 
Group four submitted the patch after coding the approach proposed by the maintainers.

Groups three, six, and ten created a struct to encapsulate the parameters and mitigate duplication 
when applying the parameterized method. Group three sent the mitigation to the maintainers and 
approved it without significant comments. In contrast, the patch from group ten was rejected by 
the same maintainer who approved the patch from group three. The maintainer justification to reject 
the patch is that group ten refactored code files that has simple context, threfore adding a layer 
of abstraction to remove a duplication deacrease the code readability without clear gains on 
the codebase quality.

Groups seven and eight used the extract method to approach the mitigations. Group eight sent the 
mitigations to the maintainers, and the code changes were approved, with minor requested changes 
in code style and smaller errors fixed. For group seven mitigation, the maintainers reject their 
patch because the code changes add layers of abstractions to the code, making it more complex and 
challenging to read. The maintainers proposed that group seven remove the duplicated functions and 
insert the duplicated code multiple times where the functions were being called. This suggestion 
may be bad practice for code quality, as it adds more complexity and responsibility to a bigger 
function; it is viewed as the best choice for removing layers of abstraction and making the 
code easier to read.

Groups nine and eleven faced similar issues to those in Mitigaiton 1 on our participant observation 
experiment, where configuration files designed as macros complicated the approach to the mitigations. 
Group nine applied the parameterized method so as not to need to know the configurations in the 
generic function, which is reasonable as the duplication approach only depends on one configuration 
value. Group eleven worked with multiple configuration values, and to refactor the duplication, 
they proposed to create a bigger macro that encapsulates the other macros, which is a clever 
solution but adds abstraction and complexity to macros, which makes the code harder to debug and read.
We do not get the feedback of the maintainers on the group eleven approach to know their opinions.

Group twelve is the student who executed the tool and passed through the process of finding 
duplications. The student found a duplication in the driver that happened across eighteen code files, 
creating a patch to remove four hundred lines of duplicated code. The student sent the mitigations 
to the maintainers, and their feedback pointed to fixing the warning in the code without arguing 
about removing the duplications.

Analyzing the students' approaches to the duplications and the maintainers' feedback, we saw that 
people without previous experience in Linux kernel development could approach duplications found 
by the ArKanjo tool and become contributors to the kernel. We observed that not all duplications 
are viewed as code with bad quality, with the maintainers analyzing the trade-off between the 
purpose of the duplicated code, code readability, and the levels of abstraction added to 
address the duplications.

\subsection{Survey Results Analysis}

We asked the teaching assistant and the students who used the ArKanjo to answer a survey with 
some questions related to the tool experience, such as installation, documentation, facility of 
use, new ideas to improve the tool, etc. Appendix \ref{app:survey} has the complete survey 
with the answers.

Related to the tool installation, neither the teaching assistant nor the students faced issues 
installing on the Debian distributions. However, the teaching assistant attempted to install it 
in the Arch Linux distribution and failed to install it on this Linux distribution. 
Regarding the tool's documentation, the teaching assistant pointed out errors that worsened his 
experience. The student asked for more examples and tutorials on using the tool in text or 
video formats.

On the ArKanjo tool preprocessing step, the student reported problems of disk space usage and 
crash errors, which we knew could happen as the tool iterates and can save information about 
every pair of functions in a codebase, as we did not implement any mechanism to control the 
memory usage. The teaching assistant reported only being able to execute the preprocessor with 
the text similarity method, and the diff method kept running for hours without finishing. 
The teaching assistant's report was unexpected behavior for us, as the text similarity method 
initially looked more complex; in contrast, the \textit{diff} method is a simple counter of equal lines 
with an operation system default command. The teaching assistant report corroborates the 
gensim efficiency \citep{gensim}.

Related to the tool's functionalities, both the student assistant and the student reported 
a missing mechanism to save the tool results on a separate file instead of the terminal output. 
The tool, by default, outputted with colors, as shown in section \ref{subsec:func}, which complicates 
redirecting the terminal output to a file.

We asked for the overall impression of the tool and if it met their expectations. We also asked 
them to give a score of one to five for installation, documentation, preprocessing, and functionalities. 
Both the teaching assistant and the student scored at least four on each mentioned point, thought that 
the tool fulfilled their expectations, and had an overall good experience. Finally, the teaching assistant 
the the student reported features ideas to the tool that may be interesting, which 
are showed below.

\paragraph{List of Possible Improvements}

\begin{itemize}

\item Enable the possibility of having several codebases of preprocessor artifacts and selecting one of them.

\item Add different methods to find duplication.

\item Adding more filters to search for duplicates, such as finding duplicates with more than 90\% 
similarity and more than 20 lines repeated.

\item Add options to highlight and filter duplications on the same file or directory.

\item Addition of support to save the tool output in files without manual preprocessing of the colored output.

\end{itemize}

With the survey feedback, we understood that users had a good experience and found value in the 
ArKanjo tool. They also provided critical points that could be improved. The most critical 
reported issues, such as documentation errors and a lack of support for multiple Linux distributions 
and different operating systems, have been fixed. Less critical issues remain open points that 
can be addressed in future work.

\subsection{Students Contribution to the Tool}

We listed multiple issues in the tool for the two groups that opted to contribute directly 
to developing the ArKanjo tool. The first group opted to work on expanding the tool's support 
for multiple operating systems. Some libraries used in the tool development only supported the 
Ubuntu distribution, and the first group was responsible for migrating these libraries to 
libraries with support for multiple operating systems. This issue was of our knowledge and 
was pointed out by the teaching assistant in the survey, as he reported the problem while 
using the tool.

The second group opted to work on a webpage for the tool documentation and support for 
light and dark modes. The colored output presented in the figures in Chapter \ref{cha:tool} was 
hard coded at first, not considering the user terminal colors, thus giving a bad user experience. 
The website helps facilitate people's contribution or use of the tool. We did not know how to 
approach the light and dark modes at first and did not need to look at how to fix the issue as 
the group searched and found a solution.

The students successfully delivered the issues proposed. We found it interesting to see 
people able to contribute to the tool on the problems that we did not even know how to approach.
The issues help the user experience and facilitate other research using the ArKanjo tool that 
we may explore in future works. Appendix \ref{app:contribution} references the 
students' contributions, the issue list proposed to the students, and figures from 
the documentation webpage.

### 2025-06-01 conteudo/tool/architecture.tex
fix grammar with grammarly

For a source file in the C programming language, we extract the code blocks with depth 0, as per the programming 
language's grammar, functions are defined in a global context at depth 0. To extract the code blocks, we firstly need to 
identify the portions of code that are comments or \textit{\#define} directives and we mark them as irrelevant, after that, 
it is simple work of finding the curly brackets that are in depth 0, and with that we find the body of the functions.

To extract the header of the functions, we start at the open curly brackets and go back in the source file consireding 
the programming language grammar  and the portions of code that have been marked as irrelevant because of comments and 
\textit{\#define} directives.

There are other elements in depth 0 that are not functions, we analyze if the code blocks is of a function at the same 
we try to extract the header of the code block.

@@

For the Java programming language, we have decided to implemente a simpler approach that does not support 
the programming language completely, as the only Java codebase explored in this work is the BigCloneBench \citep{bigclonebench} and 
the approach implemented cover the needs of this codebase.

For a source file in Java, we extract the code blocks with depth 1 and some code lines before the definition of the code block. 
Then, we check if the code blocks are functions by analyzing if the first non-empty character before the curly brackets of 
the code block is a closing parenthesis.

As per Java's grammar, functions are defined inside classes and can exist at different depths, as Java accepts inner classes, 
that means that this approach does not work with inner classes. The challenge to implement a more generic approach 
is that other elements of depth 2 or deeper, such as for-loops, can follow the closing parenthesis pattern, necessitating 
a better analysis to different these elements. The BigCloneBench \citep{bigclonebench} does not contains inner classes.

@@

We implemented two methods for code duplication detection in the tool, in which the user can choose what to use.
The first method is based in text similarity, the second is based in simpler method based on the number of equal lines.
The experiments and tests on this research were done using the text similarity method, the purpose of the 
second method is to demonstrate that the tool can have multiple duplication detection methods.

@@

For the difference method, 
we implemented a method considering the number of exactly equal lines between two functions.
For two functions \textit{function1} and \textit{function2}, let $NL_1$ be the number of lines in \textit{function1},
$NL_2$ be the number of lines in \textit{function2}, and $EQ$ be the number of equal lines between \textit{function1} and
\textit{function2}, we compute the similarity with the formula:

@@

This code duplication method is considerably simpler and more explainable then the text similarity method, 
as the metric is the ratio of common lines between the two functions. To compute the number of equal lines, we use
the \textit{diff command} built-in in the Linux environment \citep{diffcommand}.

The difference method was not tested on the evaluation of the tool and other phases of this research. 
The method was implemented to show that tool can have multiple duplication detection methods.

>>>>

For a source file in the C programming language, we extract the code blocks with depth 0; 
as per the programming language’s grammar, functions are defined in a global context at depth 0. 
To extract the code blocks, we first need to identify the portions of code that are comments or 
\textit{\#define} directives and mark them as irrelevant. After that, it is simple work to find 
the curly brackets in depth 0, and with that, we find the body of the functions.

To extract the header of the functions, we start at the open curly brackets and go back in the 
source file considering the programming language grammar and the portions of code that have been 
marked as irrelevant because of comments and \textit{\#define} directives.
There are other elements in depth zero that are not functions. We analyze whether the code blocks 
are functions at the same time that we extract the header of the code block.

@@

For the Java programming language, we have decided to implement a more straightforward approach 
that does not support the programming language completely, as the only Java codebase explored in 
this work is the BigCloneBench \citep{bigclonebench}, and the approach implemented covers the 
needs of this codebase.

For a source file in Java, we extract the code blocks with depth one and some code lines before 
defining them. Then, we check if the code blocks are functions by analyzing whether the first 
non-empty character before the curly brackets of the code block is a closing parenthesis.

As per Java’s grammar, functions are defined inside classes and can exist at different depths. 
As Java accepts inner classes, this approach does not work with inner classes. The challenge 
to implementing a more generic approach is that other elements of depth two or deeper, such 
as for-loops, can follow the closing parenthesis pattern, necessitating a better analysis of 
these elements. The BigCloneBench \citep{bigclonebench} does not contain inner classes.

@@

We implemented two methods for code duplication detection in the tool, which the user can 
choose to use. The first method is based on text similarity, and the second is simpler and 
based on the number of equal lines. The experiments and tests in this research were done 
using the text similarity method. The purpose of the second method is to demonstrate that 
the tool can have multiple duplication detection methods.

@@

For the difference method, 
we implemented a method that considered the number of exactly equal lines between two functions.
For two functions \textit{function1} and \textit{function2}, let $NL_1$ be the number of lines in \textit{function1},
$NL_2$ be the number of lines in \textit{function2}, and $EQ$ be the number of equal lines between \textit{function1} and
\textit{function2}, we compute the similarity with the formula:

@@

This code duplication method is considerably simpler and more explainable than the text similarity method, 
as the metric is the ratio of common lines between the two functions. To compute the number of equal lines, we use
the \textit{diff} command built-in in the Linux environment \citep{diffcommand}.

We did not test the Diff method during the tool evaluation and other phases of this research. 
We implemented the technique to show that a tool can have multiple duplication detection methods.

### 2025-06-01 conteudo/tool/tool.tex
add paragraph about contrib and remove some todo
fix grammar with grammarly

The tool operates in two main parts, the \textbf{Preprocessor} and the \textbf{Query Responder}. 
The \textbf{Preprocessor} is responsible to perform the heavy computations to find the code duplications 
accross a codebase and produce artifact to consulte information related to the duplications in a structured form.
The \textbf{Query Responder} consumes the artifacts produced by the Preprocessor to execute the tool functionalites as requested.
This choice enable the tool to concentrate the heavy and slows parts that can be execute only once per codebase, 
thus allowing a fast performance for multiple queries related to a single codebase.
More detailed descriptions of the Preprocessor and Query Responder can be found in
Sections \ref{subsec:architecture} and \ref{subsec:func}.

@@

\todo[inline]{Big maybes: Describe more how the functionalites are implemented, as the focus shifted for the tool as of now.}

>>>>

The tool operates in two main parts: the \textbf{Preprocessor} and the \textbf{Query Responder}. 
The \textbf{Preprocessor} is responsible for performing heavy computations to find the code 
duplications across a codebase and produce artifacts to consult information related to the 
duplications in a structured form. The \textbf{Query Responder} consumes the artifacts produced 
by the Preprocessor to execute the tool functionalities as requested. This choice enables the 
tool to concentrate on the heavy and slow parts that can be executed only once per codebase, 
thus allowing a fast performance for multiple queries related to a single codebase.

@@
